Short-time simulations of open collective quantum spin systems

### 2019-04-11 soc_squeezing/collective_paper/collective_spin.tex
minor wording changes

As a final example of collective-spin physics that is numerically
accessible via the TST expansion of Heisenberg operators, we consider
the calculation of collective-spin two-time correlation functions.  In
particular, we consider the effect of decoherence on the short-time
behavior of the two-time connected correlator
\begin{align}
  C\p{t}
  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
  \label{eq:two_time}
\end{align}
in the context of the squeezing models in Section \ref{sec:squeezing}.
In an equilibrium setting, correlation functions of this sort contain
information about the linear response of Heisenberg operators to
perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
setting, they contribute to short-time linear response (see Appendix
\ref{sec:linear_response}).  Similar correlators have made appearances
as order parameters that diagnose time-crystalline phases of
matter\cite{tucker2018shattered}.  Related multi-time correlators,
known as out-of-time-ordered correlators, are commonly examined for
signatures of quantum chaos and information
scrambling[CITE]\cite{CITE}.

\begin{figure*}
  \centering
  \includegraphics{Sp_Sm.pdf}
  \caption{Real (solid lines) and imaginary (dashed lines) parts of
    the two-time connected correlator $C\p{t}$ defined in
    Eq.~\eqref{eq:two_time} for $N=10^4$ spins initially in the
    polarized state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
    evolving under the squeezing Hamiltonians in
    Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.  Results are shown for time
    evolution without decoherence (leftmost panel), as well as with
    decoherence at rates $\gamma_n$ or $\Gamma_n$, respectively
    corresponding to single-spin or collective jump operators
    $s_n^{(j)}$ or $S_n$.  All axes in the sub-figures are shared.  At
    short times, the correlations captured by $C\p{t}$ are suppressed
    by single-spin decay ($\gamma_-$), collective decay ($\Gamma_-$),
    and single-spin dephasing ($\gamma_\z$); these correlations are
    relatively insensitive to collective dephasing ($\Gamma_\z$), and
    can change sign or grow under single-spin ($\gamma_+$) or
    collective ($\Gamma_+$) excitation.}
  \label{fig:two_time}
\end{figure*}

Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
spins, initially in the state
$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
with and without various sources of decoherence.  The two-time
correlator $C\p{t}$ degrades at short times under certain types of
decoherence, but seems relatively stable to collective dephasing (with
jump operator $S_\z$), and can qualitatively change behavior or grow
in the presence of single-spin or collective excitation (with jump
operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time} serves as
an example for the type of quantity and behavior that is accessible at
short times with the TST expansion.  This example is straightforward
to extend to multi-time and out-of-time-ordered correlators (OTOCs),
as well as equilibrium settings.

>>>>

As a final example of collective-spin physics that is numerically
accessible via the TST expansion of Heisenberg operators, we consider
the calculation of collective-spin two-time correlation functions.  In
particular, we consider the effect of decoherence on the short-time
behavior of the two-time connected correlator
\begin{align}
  C\p{t}
  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
  \label{eq:two_time}
\end{align}
in the context of the squeezing models in Section \ref{sec:squeezing}.
In an equilibrium setting, correlation functions of this sort contain
information about the linear response of Heisenberg operators to
perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
setting, they contribute to short-time linear response (see Appendix
\ref{sec:linear_response}).  Similar correlators have made appearances
as order parameters that diagnose time-crystalline phases of
matter\cite{tucker2018shattered}, and related multi-time correlators,
known as out-of-time-ordered correlators, are commonly examined for
signatures of quantum chaos and information
scrambling[CITE]\cite{CITE}.

\begin{figure*}
  \centering
  \includegraphics{Sp_Sm.pdf}
  \caption{Real (solid lines) and imaginary (dashed lines) parts of
    the two-time connected correlator $C\p{t}$ defined in
    Eq.~\eqref{eq:two_time} for $N=10^4$ spins initially in the
    polarized state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
    evolving under the squeezing Hamiltonians in
    Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.  Results are shown for time
    evolution without decoherence (leftmost panel), as well as with
    decoherence at rates $\gamma_n$ or $\Gamma_n$, respectively
    corresponding to single-spin or collective jump operators
    $s_n^{(j)}$ or $S_n$.  All axes in the sub-figures are shared.  At
    short times, the correlations captured by $C\p{t}$ are suppressed
    by single-spin decay ($\gamma_-$), collective decay ($\Gamma_-$),
    and single-spin dephasing ($\gamma_\z$); these correlations are
    relatively insensitive to collective dephasing ($\Gamma_\z$), and
    can change sign or grow from spontaneous single-spin ($\gamma_+$)
    or collective ($\Gamma_+$) excitation.}
  \label{fig:two_time}
\end{figure*}

Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
spins, initially in the state
$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
with and without various sources of decoherence.  The two-time
correlator $C\p{t}$ degrades at short times under certain types of
decoherence, but seems relatively insensitive to collective dephasing
(with jump operator $S_\z$), and can qualitatively change behavior or
grow in the presence of single-spin or collective excitation (with
jump operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time}
serves as an example for the type of quantity and behavior that is
accessible at short times with the TST expansion.  This example is
straightforward to extend to multi-time and out-of-time-ordered
correlators (OTOCs), as well as equilibrium settings.
